---
title: "Scratch pad for Colinearity Analysis"
author: "Javier Matias-Cabrera, javierm4"
date: ''
output:
  html_document:
    toc: yes
  pdf_document: default
urlcolor: cyan
---

```{r setup, include=FALSE}
library(knitr)

#knitr::opts_chunk$set(echo = TRUE)
#options(digits=8)
knit_hooks$set(htmlcap = function(before, options, envir) {
  if(!before) {
      paste('<p style="text-align: center;" class="caption">',options$htmlcap,"</p>",sep="")
    }
})
```

```{r message=FALSE, warning=FALSE}
library(jtools)

source("load_dataset.r")
source("transformations.r")
source("plots.r")

```


### Introduction

For this study, we will perform some exploratory data analysis to explain how the effect of a continous predictor on the response varies across different treatment groups (categorical predictor). Namely, we'll test:


$H_0: \beta_{j,LivArea:Neighborhood} = 0, \forall j \text{ - The mean price per square foot of a house is the same for every neighborhood j }$

$H_1: \beta_{j,LivArea:Neighborhood} \neq 0,  \forall j \text{ - At least one neighborhood j has a different mean price per square foot }$

We will also attempt to infer which neighborhoods place the highest monetary value on living area and determine whether the living area is collinear with other predictors.

### Methods

We will begin by loading the housing data stored in `AmesHousing.csv`. We will also define a convenience function `splitDataset` that will randomly split the data into two data sets: *"train"* and *"test"*.
```{r message=FALSE, warning=FALSE}

splitDataset = function(data) {
  training = createDataPartition(data$SalePrice, times=1, p=0.80, list=FALSE)
  return(list(train=data[training,], test=data[-training,]))
}

housing_data = read_dataset("../data/AmesHousing.csv")

splits = splitDataset(housing_data)
training_data = splits$train
test_data = splits$test
```

Now we can begin our exploratory data analysis. To examine the relationship between `Neighborhood` and `SalePrice`, we will perform a simple one-way ANOVA to determine whether belonging to different neighborhoods has any effect on `SalePrice` at all.

```{r}
neighborhood_oneway_anova = aov(SalePrice ~ Neighborhood, data=housing_data)
neigh_oneway_anova_summary = summary(neighborhood_oneway_anova)
```

A simple linear model would allow us to infer whether the General Living Area (`Gr.Liv.Area`) of a house has any relationship with `SalePrice`.

```{r}
livarea_model = lm(SalePrice ~ Gr.Liv.Area, data=housing_data)
livarea_model_summary = summary(livarea_model)
```

And, in order to infer whether or not a house's living area is more expensive in some neighborhoods than others, we'll perform Analysis of Co-Variance (ANCOVA) by including an interaction term between the continous variable `Gr.Liv.Area` and the categorical variable `Neighborhood`. If the slopes of `Gr.Liv.Area` are different (not parallel) for different neighborhoods, then we would notice a strong interaction between the two predictors.
```{r}
neigh_livarea_interaction_model = lm(SalePrice ~ Gr.Liv.Area*Neighborhood, data=housing_data)
neigh_livarea_anova = anova(neigh_livarea_interaction_model)
p_value_neigh_livarea_interact =anova(neigh_livarea_interaction_model)[3,5]
```

We will also estimate actual slopes of the lines of `SalePrice` vs. `Gr.Liv.Area` for different neighborhoods to examine any differences.
```{r}
livarea_slopes_raw = sim_slopes(neigh_livarea_interaction_model, pred=Gr.Liv.Area, modx=Neighborhood,
                        johnson_neyman = FALSE,centered= "none", confint=T)

livarea_slopes = data.frame(livarea_slopes_raw$slopes[,1:5],stringsAsFactors=FALSE)
livarea_slopes[,2:5] = sapply(livarea_slopes[,2:5], as.double)
livarea_slopes[,2:5] = sapply(livarea_slopes[,2:5], signif, digits=4)

colnames(livarea_slopes)[1] = "Neighborhood"
colnames(livarea_slopes)[3] = "Std. Err."
colnames(livarea_slopes)[4] = "2.5%"
colnames(livarea_slopes)[5] = "97.5%"
```


Lastly, in order to determine whether the effect of `Gr.Liv.Area` on `SalePrice` is not explained by other predictors, we'll calculate the partial correlation coefficient between them with the effect of all other predictors removed. 




### Results

#### Neighborhood Effect on Living Area Price


From out first ANOVA of `Neighborhood` and `SalePrice`, it can easily be seen that the neighborhood had a significant effect on the price. We can also see this graphically using a box-plot.


```{r fig.height=5, fig.width=10}
neigh_oneway_anova_summary
boxplot(SalePrice ~ Neighborhood, data = housing_data,las=2, col=2:8, pch=20, cex=.7,
        main="Sale Prices by Neighborhood",
        ylab="Sale Price (in US Dollars")
```

Similarly, it is easily shown that `Gr.Liv.Area` has an effect on `SalePrice`.

```{r fig.height=5, fig.width=10}
plot(SalePrice ~ Gr.Liv.Area, data = housing_data,
     xlab = "Living Area (square feet)",
     ylab = "Sale Price (US Dollars)",
     main = "Sale Price vs General Living Area",
     pch  = 20,
     cex  = .5,
     col  = "grey")
abline(livarea_model, col="darkorange", lwd=2)
```


However, we also observe that the interaction model between `Neighborhood` and `Gr.Liv.Area` yielded a near-zero p-value of `r signif(p_value_neigh_livarea_interact,4)`. This **suggests a significant interaction between the two predictors** when predicting `SalePrice`.

```{r}
signif(p_value_neigh_livarea_interact,4)
print(neigh_livarea_anova[-4,], signif.stars=FALSE)
```

Because of this, we obtain different slopes for `Gr.Liv.Area` depending on `Neighborhood`. Each of the slope estimates are the sum of the `Gr.Liv.Area` coefficient and its interaction term with the corresponding `Neighborhood` factor dummy variable. As with all estimates, we can calculate a mean and a confidence interval.

```{r fig.height=5, fig.width=10, htmlcap="Figure X: FOOOBAR"}

## Get the top 5 neighborhoods with the most houses
ordered_neigh = order(-table(housing_data$Neighborhood))

interact_plot(neigh_livarea_interaction_model, pred=Gr.Liv.Area,
              modx=Neighborhood,color.class="Qual1",
              modxvals = names(table(housing_data$Neighborhood)[ordered_neigh][1:6]),
              x.label="General Living Area (square feet)",
              y.label="Sale Price (US Dollars)") +
  ggtitle("Gr.Liv.Area Slopes of the 5 Neighborhoods with Most Sales" ) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r htmlcap="Table X: FOOBAR"}
kable(livarea_slopes)
```


### Discussion

#### Relationship between Neighborhood, Sale Price and Other Predictors

Having observed how both `Neighborhood` and `Gr.Liv.Area` interact with each other and influence `SalePrice`, it is important to highlight that it **does not imply a casual relationship** between them. This becomes more evident when we examine the partial correlation coefficient of `Gr.Liv.Area` and `SalePrice` with the effect of all other 78 predictors removed. We observe that its magnitude is less than 0.02.

```{r}
without_livarea       = lm(SalePrice ~ . -Gr.Liv.Area , data=housing_data)
livarea_collinearity  = lm(Gr.Liv.Area~ . -SalePrice, data=housing_data)
(partial_corr_livarea_saleprice = cor(resid(without_livarea),resid(livarea_collinearity)))
```
 
Therefore, general living area is collinear with many of the other predictors available in the dataset, and does not independently predict price. We can, however, **reject the null hypothesis** that the average price per square foot of living space is equal for all neighborhoods with a confidence level $\alpha= 0.01$.



